Cell-type diversity and connectome of the ctenophore apical organ

### 2024-03-08 manuscript/Manuscript.qmd
MS: add refs, update bib

---
title: "New paper template - title here"
mainfont: Arial
monofont: DejaVu Sans Mono #Times New Roman
fontsize: 11pt
linestretch: 1.2
linkcolor: blue
documentclass: article
max-width: 100%
number-sections: false
format:
  html: default
  pdf:
    pdf-engine: xelatex
    number-sections: true
    margin-left: 0.8in
    margin-right: 0.8in
    margin-top: 0.8in
    margin-bottom: 0.8in
    toc: true
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{lineno}
    include-before-body:
      text: |
        \linenumbers
bibliography: references.bib #add bibliography in this file
csl: elife.csl #change according to journal
---

@@

Author 1^1,2,\*^ , Author 2^2^, Author 3^1,2^\

^1^Heidelberg University, Centre for Organismal Studies (COS), 69120 Heidelberg, Germany\
^2^Living Systems Institute, University of Exeter, Exeter, EX4 4QD, United Kingdom\
^\*^Correspondence: corr_author\@email.com

@@

You can add references either by referring to their id in the .bib file e.g., [@Marinkovi2020], or by switching to the visual editor (Cogwheel in the .Rmd menu -\> Use Visual Editor). The awesome paper by [@Jokura_2019].

In the visual editor mode, go to 'Insert' -\> \@ Citation

You can select a Zotero library, PubMed, CrossRef etc. and insert the citations.

@@

curl -LH "Accept: application/x-bibtex" https://doi.org/10.7554/eLife.91258.1 >> references.bib
```

*Platynereis dumerilii* is a marine annelid... [@ozpolat2021]

The references are stored in manuscript/references.bib (need to be defined in the Yaml header). This file will automatically updated when you insert a new reference through the Visual editor \> Insert \> Citations.

In this documents, references will be formatted in the style of eLife. This is defined in the Yaml header under: csl: elife.csl. The elife.csl file is saved in the /manuscript folder.

If you would like to use a different citation format, download the respective .csl file (e.g., from the Zotero style repository <https://www.zotero.org/styles>), save it in the /manuscript folder of the project and change the Yaml to csl: your_favourite_journal.csl.

@@

@@

We would like to thank the Jekely lab for the R project template (<https://github.com/JekelyLab/new_paper_template>) we used to write this paper. This work was funded by ...

>>>>

---
title: "Cell-type diversity and connectome of the ctenophore apical organ"
mainfont: Arial
monofont: DejaVu Sans Mono #Times New Roman
fontsize: 11pt
linestretch: 1.2
linkcolor: blue
documentclass: article
max-width: 100%
format:
  html: default
  pdf:
    pdf-engine: xelatex
    number-sections: false
    margin-left: 0.8in
    margin-right: 0.8in
    margin-top: 0.8in
    margin-bottom: 0.8in
    toc: false
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{lineno}
    include-before-body:
      text: |
        \linenumbers
bibliography: references.bib #add bibliography in this file
csl: elife.csl #change according to journal
---

@@

Kei Jokura^1,3,\*^, Sanja Jasek^1,2^, Lara Kewalow^2^, Pawel Burkhardt^4^, Gáspár Jékely^1,2,\*^\

^1^Living Systems Institute, University of Exeter, Exeter, EX4 4QD, United Kingdom\
^2^Heidelberg University, Centre for Organismal Studies (COS), 69120 Heidelberg, Germany\
^3^National Institute for Basic Biology, Okazaki, ..., Japan
^4^Sars Centre ... Berge, Norway
^\*^Correspondence: corr_author\@email.com

@@

First descriptions of AO: [@aronova1974electron]

The lobate ctenophore *Mnemiopsis*

The lithocytes have a large calcium carbonate concretion and... [@Tamm_2014]
Living lithocytes are transported along the balancer cilia to the tip of the statocyst [@Noda_2014]
The projections of a group of bridge cells located on two sides of the apical organ forms an arch and connects the balancers along the tentacular plane [@Tamm_2002].

@@

#curl -LH "Accept: application/x-bibtex" https://doi.org/10.7554/eLife.91258.1 >> references.bib
```

@@

### Volume EM reconstruction of the *Mnemiopsis* apical organ

- specimen 
- volume, resolution, fixation, quality
- number of sections
- skeletonisation in Catmaid
- number of cells
- general anatomy, regions, quadrants

Figure1: (pic of a larva, boxed region of AO, schematic, catmaid pic with slice, dimensions, quadrants, regions)

### Classification of cell types

- ultrastructural criteria (cilia, vesicles etc)
- number of types, cells per type - table

Figure2 - Balancer complex: balancer, lithocyte, groove 2 types, bridge - EM pics

"the larval bridge has not yet developed forks at its ends, as seen in adult Mnemiopsis bridges"[@Tamm_2002]

Figure3 - dome, bicil, monocil, no-cil, plumose, dense vesicle cells (rename), gland? cells, oligo/multiciliated cells

Plumose cells [@Hernandez_Nicaise_1984] may be pressure sensors [@Tamm_2014b]

Figure4 - lamellate body complex, LB, intra-multi-ciliated, relationship to balancer

Figure5 - nerve net, different fragments/cells/syntytia, EM, relation to entire organ

Figure 6 -  cilia? renderings of basal bodies, centrioles, length, axoneme structure (table), table about cell types, number of cilia etc.

## Synapses

"synaptic regions with the characteristic presynaptic triad mor-
phology of ctenophore nerves (i.e., a single layer of
vesicles, smooth ER sac, and closely apposed mito-
chondria; Horridge and Mackay, 1964; Hernandez-
Nicaise, 1973)"

Figure 7 - synapses, reconstruction, annotation, numbers, distribution of partners, example EM (figure supplement - lots of EM), mitochondria
(Fig suppl - mitochondrion stats - Sanja)

Gap juntions? Innexins are expressed in the apical organ in specific patches of cells in each quadrant [@Ortiz_2023]

Ctenophore synapses [@Hernandez_Nicaise_1973]

We found no synapses between bridge and balancer, in agreement with Tamm [@Tamm_2002].

Figure 8 - connectome
cell-type based, quadrant based, nerve net syntycia, bridge, quadrants
(Fig suppl - cell-based network)

Tamm&Tamm observed synapses from neurites to bridge cells (2002).
and in Mnemiopsis larvae observed synapses from bridge cells onto other cells

@@

We would like to thank the Jekely lab for the R project template (<https://github.com/JekelyLab/new_paper_template>) we used to write this paper. This work was funded by ...
JSPS...
ERC..
Others ...
